Parameter estimation

### 2023-03-10 SFE-article/Article_Model_Parameters.tex
10.03

@@

The quasi-one-dimensional compressible Navier-Stokes equations in Cartesian coordinates is given be equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_2}.

{\footnotesize
\begin{align}
		\label{EQ: CompressibleEuler_1}
        \cfrac{\partial \left( \rho_f A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v \right)}{\partial z} &= 0 \\
        \cfrac{\partial \left( \rho_f v A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v^2 \right)}{\partial z} &= -A_f \cfrac{\partial P}{\partial z} \\
        \cfrac{\partial \left( \rho_f e A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v e\right)}{\partial z} &= -P\cfrac{\left( A_f v \right)}{\partial z} + \cfrac{\partial}{\partial z} \left( \cfrac{\partial T}{\partial z} \right)   
        \label{EQ: CompressibleEuler_2}
    \end{align}  
}

@@

\section{Results} \label{CH: Results}

@@

%\subsection{Bayes theorem} \label{CH: Bayes}
%\subfile{Sections/Bayes_Theorem}

>>>>

\usepackage[final]{pdfpages}

@@

The quasi-one-dimensional compressible Navier-Stokes equations in Cartesian coordinates is given be equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_3}.

{\footnotesize
\begin{align}
		\label{EQ: CompressibleEuler_1}
        \cfrac{\partial \left( \rho_f A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v \right)}{\partial z} &= 0 \\
        \cfrac{\partial \left( \rho_f v A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v^2 \right)}{\partial z} &= -A_f \cfrac{\partial P}{\partial z} {EQ: CompressibleEuler_2} \\
        \cfrac{\partial \left( \rho_f e A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v e\right)}{\partial z} &= -P\cfrac{\left( A_f v \right)}{\partial z} + \cfrac{\partial}{\partial z} \left( \cfrac{\partial T}{\partial z} \right)   
        \label{EQ: CompressibleEuler_3}
    \end{align}  
}

@@

\subsection{Experimental work}
\subfile{Sections/Experiments}

\section{Results}
\subfile{Sections/Results}

@@

\subsection{Bayes theorem} \label{CH: Bayes}
\subfile{Sections/Bayes_Theorem}

\clearpage
\onecolumn
\subsection{Solid density measurement} \label{CH: Solid_Density_Measurment}

\begin{figure*}[b!]
	%\centering 
	\includegraphics[trim=0cm 6cm 0cm 0cm, clip,height=0.9\textheight]{Sections/ultraReportT5.pdf}
	\caption{The result of solid density measurement}
\end{figure*}

### 2023-03-10 SFE-article/Sections/Model.tex
10.03

        		\begin{axis}[
        			legend style={draw=none},
        			xmin = 0, xmax = 10,
        			ymin = -0.5, ymax = 1.5,
        			xlabel = {$c_f[kg/m^3]$},
        			ylabel = {$\Gamma(c_f)[-]$}]
        			\addplot[
        			domain = 0:10,
        			samples = 100,
        			smooth,
        			thick,
        			blue,
        			] { 1 / (1 + exp( 2*5 .* ( x - 5 ) ) ) };
        			\addplot[thick, samples=50, dashed, domain=0:1.5, black] coordinates {(5,-0.5)(5,1.5)};

@@

        {\color{blue} The heat equation was introduced in the previous chapter. The heat balance describe movement of the internal energy in the system. For real gases it is complicated to write the heat balance in terms of temperature. Alternatively, the temperature can be obtained based on the equation of state if the internal energy and pressure are know. A roootfinder can be used to find a value of temperature, which satisfy the equation of state given values of internal energy and pressure. The temperature needs to be reconstructed from the internal energy in every time-step.
		}

>>>>

        		\begin{axis}[
        			legend style={draw=none},
        			xmin = 0, xmax = 10,
        			ymin = -0.5, ymax = 1.5,
        			xlabel = {$c_f[kg/m^3]$},
        			ylabel = {$\gamma(c_f)[-]$}]
        			\addplot[
        			domain = 0:10,
        			samples = 100,
        			smooth,
        			thick,
        			blue,
        			] { 1 / (1 + exp( 2*5 .* ( x - 5 ) ) ) };
        			\addplot[thick, samples=50, dashed, domain=0:1.5, black] coordinates {(5,-0.5)(5,1.5)};

@@

		\todo[inline]{Draft of new heat equation starts here. TODO: \\ Check if the heat cpacity of solid phase needs to be added \\ Elaborate why pressure is know}

		\noindent\colorbox{pink}
		{
		\begin{minipage}{\columnwidth}
			The heat equation was introduced in the previous chapter through equation \ref{EQ: CompressibleEuler_3} or

			{\footnotesize
			\begin{equation*}
				\cfrac{\partial \left( \rho_f e A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v e\right)}{\partial z} = -P\cfrac{\left( A_f v \right)}{\partial z} + \cfrac{\partial}{\partial z} \left( \cfrac{\partial T}{\partial z} \right) 
			\end{equation*}
			}

			For real gases the internal energy is defined as

			{\footnotesize
				\begin{equation}
					de=C_vdT-\left[P-T\left(\cfrac{\partial P}{\partial T}\right)_{V_m}\right]dv_m
					\label{EQ: internal_energy_def}
				\end{equation} }

			If a gas is consider to be perfectly caloric ($e=C_v T$), then the energy equation can be written in form of temperature. The perfectly caloric gas can be seen as the special case of a real gas, where the second term of the equation \ref{EQ: internal_energy_def} goes to zero.

			For real gases it is complicated to write the heat balance in terms of temperature. Alternatively, the energy equation can be used in the form of internal energy, so as it is given by equation \ref{EQ: CompressibleEuler_3}. In such a case, the temperature (which is used as the input to some other functions) needs to be recover from the internal energy. A relation for the internal energy can be obtained from an equation of state. For Peng-Robinson, such a relation is given by equation \ref{EQ: internal_energy_PR}

			{\footnotesize
				\begin{equation}
					e = C_vT + \cfrac{a\left(\alpha -T\cfrac{d \alpha}{dT}\right)}{2\sqrt{2}b} \ln \left[ \cfrac{1+b\left(1-\sqrt{2}\rho\right) }{1+b\left(1+\sqrt{2}\rho\right)} \right]
					\label{EQ: internal_energy_PR}
				\end{equation}
			}

			To solve equation \ref{EQ: internal_energy_PR}, values of temperature, pressure and density needs to be know. If an equation of state is introduce, then only 2 out 3 variables needs to be know as the third one can be calculated, this can be represented as follow

			{\footnotesize
			\begin{equation*}
				e(t,z)=e(T(t,z),P(t),\rho(t,z))=e(T(t,z),P(t),\underbrace{\rho(T(t,z),P(t))}_{E.O.S.}) 
			\end{equation*}
			}

			If the value of internal energy $e$ is know from time evolution of the energy equation (\ref{EQ: CompressibleEuler_3}), and pressure is know from measurement, then the temperature can be reconstructed. A roootfinder can be used to find a value of temperature, which minimize the difference between value of internal energy coming from the time evolution and the output from equation \ref{EQ: internal_energy_PR}. Such a procedure allow to find local temperature along spatial direction $z$ and needs to be repeated every time-step.

		\end{minipage} }

		\todo[inline]{Add short section about boundary conditions and comment on the used numerical methods}

### 2023-03-10 SFE-article/Sections/Parameter_estimation.tex
10.03

\subsection{Least squares estimation}
A least squares parameter estimation does not require prior knowledge of the distribution of unobservable errors, yield  unbiased estimates, and results in the minimum variance among all linear unbiased estimators. If the observations $ Y$ for the model responses are continuous functions of time from $t = 0$ to $t = t_f$, the Markov (or "rigorous least squares") criterion is to minimize:

@@

\iffalse
\subsection{Weighted sum of squares - Cumulative data}

@@

\subsection{Maximum likelihood estimation}

@@

By collecting a values of $y$ and selecting the values of $\theta$ that maximize the likelihood function $L\left( \theta | y \right)$, a function described in chapter \ref{CH: Bayes} in connection with Bayes' theorem\todo{Is the reference it previous chapter needed?}. Such estimators, $\hat{\theta}$, are known as maximum likelihood estimators. In effect, the methods selects those values of $\theta$ that are at least as likely to generate the observed sample as any other set of values of the parameters if the probability density of $y$ were to be extensively simulated through use of the probability density $p\left( y|\theta \right)$. In making a maximum likelihood estimate, we assume that the form of the probability density (\sout{only the $\theta$ need be determined}) and that all possible values of $\theta$ are equally likely before experimentation. \todo{Needed?}

@@

		{\footnotesize
			\begin{equation}
				\begin{aligned}
					&\hat{\theta}_{MLE} &= \arg \max_{\sigma, \theta \in \Theta} \ln L = \arg \max_{\sigma,\theta \in \Theta} p(\theta|y) \\
					&\text{subject to}
					& \dot{x} = f(t,x,\theta) \\
					&& \dot{\theta} = 0 \\
					&& y = y(x)
				\end{aligned}
		\end{equation} }

@@

		{\color{red}Based on the first order optimality condition,} the $\ln L$ can be maximized with respect to the vector $\theta$ by equating to zero the partial derivatives of $\ln L$ with respect to each of the parameters:

>>>>

\subsubsection{Least squares estimation}
A least squares parameter estimation does not require prior knowledge of the distribution of unobservable errors, yield  unbiased estimates, and results in the minimum variance among all linear unbiased estimators. If the observations $ Y$ for the model responses are continuous functions of time from $t = 0$ to $t = t_f$, the Markov (or "rigorous least squares") criterion is to minimize:

@@

\iffalse
\subsubsection{Weighted sum of squares - Cumulative data}

@@

\subsubsection{Maximum likelihood estimation}

@@

By collecting a values of $y$ and selecting the values of $\theta$ that maximize the likelihood function $L\left( \theta | y \right)$, a function described in chapter \ref{CH: Bayes} in connection with Bayes' theorem\todo{Is the reference it previous chapter needed?}. Such estimators, $\hat{\theta}$, are known as maximum likelihood estimators. In effect, the methods selects those values of $\theta$ that are at least as likely to generate the observed sample as any other set of values of the parameters if the probability density of $y$ were to be extensively simulated through use of the probability density $p\left( y|\theta \right)$. In making a maximum likelihood estimate, we assume the form of the probability density and that all possible values of $\theta$ are equally likely before experimentation. \todo{Needed?}

@@

		{\footnotesize
			\begin{equation}
				\begin{aligned} \label{EQ: Optimization_formulation_MLE}
					&\hat{\theta}_{MLE} &= \arg \max_{\sigma, \theta \in \Theta} \ln L = \arg \max_{\sigma,\theta \in \Theta} p(\theta|y) \\
					&\text{subject to}
					& \dot{x} = f(t,x,\theta) \\
					&& \dot{\theta} = 0 \\
					&& y = y(x)
				\end{aligned}
		\end{equation} }

@@

		Based on the first order optimality condition, the $\ln L$ can be maximized with respect to the vector $\theta$ by equating to zero the partial derivatives of $\ln L$ with respect to each of the parameters:

### 2023-03-10 SFE-article/Sections/Thermo.tex
10.03

    where $A$ and $B$ are parameters as defined in the appendix. The roots of the polynomial can be found iteratively or by Cardano formula. Depends on the operating conditions, one or two roots can be found. In one-phase region the fluid can be describe as gas, liquid or super-critical. In the two-phase region the gas-liquid mixture is present. The biggest root is assigned to the gas phase and smallest root corresponds to the liquid phase.

@@

        \begin{figure*}[H]
			\centering
			\includegraphics[trim = 1.5cm 4.0cm 2.5cm 10.0cm,clip,width=\textwidth]{Figures/MU.pdf}	
			\caption{Viscosity obtained based on different correlations ({\color{red}Decrease space between figures})} 
            \label{fig: SFE_Properties_mu}
		\end{figure*}

>>>>

    where $A$ and $B$ are parameters as defined in the appendix (\ref{CH: Thermodynamic_details}). The roots of the polynomial can be found iteratively or by Cardano formula. Depends on the operating conditions, one or two roots can be found. In one-phase region the fluid can be describe as gas, liquid or super-critical. In the two-phase region the gas-liquid mixture is present. The biggest root is assigned to the gas phase and smallest root corresponds to the liquid phase.

@@

        \begin{figure*}[H]
			\centering
			\includegraphics[trim = 1.5cm 11.5cm 2.5cm 10.0cm,clip,width=\textwidth]{Figures/MU.pdf}	
			\includegraphics[trim = 1.5cm 4.0cm 2.5cm 19.0cm,clip,width=\textwidth]{Figures/MU.pdf}	
			\caption{Viscosity obtained based on different correlations } 
            \label{fig: SFE_Properties_mu}
		\end{figure*}
